---
title: "GAPMINDER | LIFE EXPECTANCIES 2007"
format: dashboard
theme: Lux
---

```{r}
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, reactable, bslib, bsicons, shiny, reactable, gapminder, rnaturalearth, plotly, countrycode, htmltools)

```

```{r}

gap_07 <- gapminder %>% filter(year == 2007)

highest_life_exp <- gap_07 %>% 
  arrange(-lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)

lowest_life_exp <- gap_07 %>% 
  arrange(lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)

mean_life_exp <- gap_07 %>% 
  summarise(weighted_mean = sum(lifeExp * pop)/ sum(pop)) %>% 
  round(1)

```

```{r}

country_shapes <- rnaturalearth::ne_countries()

gap_07_mid <- gap_07 %>% 
  mutate(country_code = countrycode(country,
                             "country.name",
                             "iso3c"))


gap_07_shapes <- left_join(gap_07_mid, country_shapes, by = c("country_code" = "adm0_a3"))


gap_07_shapes <- gap_07_shapes %>% mutate(
  tooltip_label = paste(country, 
                        round(lifeExp, 1),
                        sep = ": ")
)

gap_map <- ggplot(data = gap_07_shapes) + 
  geom_sf(aes(geometry = geometry, fill = lifeExp, text = tooltip_label)) +
  theme_void() +
  theme(legend.position = "none")

#gap_plotly <- ggplotly(gap_map, tooltip = "text")

```


```{r}

top_20 <- gap_07 %>% arrange(desc(lifeExp)) %>% 
  head(20) %>% 
  ggplot(aes(y = reorder(country, lifeExp), 
             x = lifeExp,
             fill = lifeExp)) +
  geom_col() +
  geom_text(aes(label = round(lifeExp, 1)),
            nudge_x = -10,
            color = "white"
            ) +
  labs(y = "Country", 
       x = "Life Exp") +
  theme(legend.position = "none")
```


# HOME

## Row 1 {height=20%}

### Column 1
```{r}
value_box(
  title = "Highest Life Expectancy",
  value = highest_life_exp,
  showcase = bsicons::bs_icon("heart"),
  theme = "primary"
)
```

### Column 2

```{r}
value_box(
  title = "Lowest Life Expectancy",
  value = lowest_life_exp,
  showcase = bsicons::bs_icon("thermometer-low"),
  theme = "primary"
)
```

### Column 3

```{r}
value_box(
  title = "Average Life Expectancy",
  value = mean_life_exp,
  showcase = bsicons::bs_icon("front"),
  theme = "primary"
)
```

## Row 2 {height=80%}

### Column 1 {width=70%}
```{r title = "Map of countries by life expectancy" } 

gap_map

```

### Column 2
```{r}
"Top 20 countries by life expectancy"

top_20

```


# Download Data

```{r}

library(htmltools)

htmltools::browsable(
  tagList(
    reactable(
      gapminder, defaultPageSize = 5,
      elementId = "gapminder-table",
      searchable = T,
      filterable = T),

tags$button("Download as CSV",
            onclick = "Reactable.downloadDataCSV('gapminder-table')")
      
    )
  )


```


# About

Our mission is to fight devastating ignorance with a fact-based worldview everyone can understand.

Gapminder identifies systematic misconceptions about important global trends and proportions and uses reliable data to develop easy to understand teaching materials to rid people of their misconceptions.

Gapminder is an independent Swedish foundation with no political, religious, or economic affiliations.

<iframe width="560" height="315" src="https://www.youtube.com/embed/v7WUpgPZzpI?si=a-KiwskbFBQmNCX6" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>


